Locality and Effect Reflection

### 2025-09-17 paper.tex
Add skeleton for more complete appendices

\newcommand{\nat}{\mathbb{N}}
\newcommand{\glob}{\mathop{\Box}}
\newcommand{\cat}{\mathbb{C}}
\newcommand{\catobj}{C}
\newcommand{\op}{\mathtt{op}}
\newcommand{\catop}{\cat^{\op}}
\newcommand{\id}[1]{\mathrm{id}_{#1}}
\newcommand{\setc}{\mathrm{Set}}
\newcommand{\psh}{\widehat{\cat}}
\newcommand{\terminal}{\top}
\newcommand{\initial}{\bot}
\newcommand{\lawc}{\mathrm{Law}}
\newcommand{\lawcop}{\lawc^{\mathtt{op}}}
\newcommand{\pshlawcop}{\widehat{\lawcop}}
\newcommand{\freef}{Free}
\newcommand{\constfname}{\Omega}
\newcommand{\constf}[1]{\constfname #1}
\newcommand{\yonedaname}{y}
\newcommand{\yoneda}[1]{y(#1)}
\newcommand{\homset}[3]{\mathrm{Hom}_{#1}(#2, #3)}
\newcommand{\arrow}[2]{{#2}^{#1}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\reflectname}{\Uparrow}
\newcommand{\reflectraw}[1]{\mathop{\reflectname_{#1}}}
\newcommand{\reflect}[3]{\reflectraw{#1}(#2, #3)}
\newcommand{\reflectc}[2]{\mathop{\reflectname_{#1 #2}}}
\newcommand{\performraw}[2]{\mathop{\reflectname(#1(#2))}}
\newcommand{\perform}[5]{\performraw{#1}{#2}(#3, #4. #5)}
\newcommand{\reifyname}{\Downarrow}
\newcommand{\reifyraw}[1]{\mathop{\reifyname_{#1}}}
\newcommand{\reify}[3]{\reifyraw{#1}(#2.\,#3)}
\newcommand{\reifyc}[2]{\mathop{\reifyname_{#1 #2}}}
\newcommand{\outlname}{\mathrm{out}_L}
\newcommand{\outl}[1]{\mathop{\outlname} #1}
\newcommand{\outrname}{\mathrm{out}_R}
\newcommand{\outr}[1]{\mathop{\outrname} #1}
\newcommand{\mon}{\mathcal{T}}
\newcommand{\ind}[1]{W_{#1}}
\newcommand{\indintro}[4]{#2_{#1}(#3,\, #4)}
\newcommand{\retname}{\mathrm{ret}}
\newcommand{\ret}[2]{\mathop{\retname_{#1}}(#2)}
\newcommand{\indelim}[3]{\mathop{\mathrm{rec}_{#1}} #2 \,\{#3\}}
\newcommand{\indretcase}[2]{\retname(#1) \Rightarrow #2}
\newcommand{\indcase}[4]{#1(#2, #3) \Rightarrow #4}
\newcommand{\seq}{\,;\,}
\newcommand{\fst}{\mathop{\pi_1}}
\newcommand{\snd}{\mathop{\pi_2}}
\newcommand{\inl}[1]{\mathop{\mathrm{in}_L} #1}
\newcommand{\inr}[1]{\mathop{\mathrm{in}_R} #1}
\newcommand{\case}[5]{\mathop{\mathrm{case}} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\abs}[3]{\mathop{\lambda}(#1 \types #2).\,#3}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\types}{\mathrel{:}}
\newcommand{\gtypes}{\mathrel{:_{\Box}}}
\newcommand{\cempty}{\varepsilon}
\newcommand{\ccons}[2]{#1;\,#2}
\newcommand{\lbind}[3]{\ccons{#1}{#2\types#3}}
\newcommand{\gbind}[3]{\ccons{#1}{#2\gtypes#3}}
\newcommand{\cdiv}[1]{#1/\Box}
\newcommand{\bdiv}[1]{\cdiv{\mathop{!_{#1}}}}
\newcommand{\return}[1]{\mathop{\mathrm{return}} #1}
\newcommand{\letv}[3]{\mathop{\mathrm{let}} #1 = #2 \mathop{\mathrm{in}} #3}
\newcommand{\gbox}[1]{\mathop{\mathrm{box}} #1}
\newcommand{\gunbox}[3]{\mathop{\mathrm{let}} (\gbox #1) = #2 \mathop{\mathrm{in}} #3}
\newcommand{\grun}[1]{\mathop{\mathrm{run}}(#1)}
\newcommand{\sbst}[3]{#1[#2\backslash#3]}
\newcommand{\turnv}{\mathrel{\vdash_V}}
\newcommand{\turnc}{\mathrel{\vdash_C}}
\newcommand{\evmap}{\mathop{ev}}
\newcommand{\currymap}{\mathop{\lambda}}
\newcommand{\reducestov}{\rightsquigarrow_V}
\newcommand{\reducestoc}{\rightsquigarrow_C}
\newcommand{\mif}{\mathop{\mathrm{if}}}
\newcommand{\mthen}{\mathop{\mathrm{then}}}
\newcommand{\melse}{\mathop{\mathrm{else}}}
\newcommand{\embed}[1]{\mathopen{\lbag}#1\mathclose{\rbag}}

@@

We define the operational semantics as a pair of mutually defined
reduction relations $\reducestov$ and $\reducestoc$ for values and
computations respectively shown in Fig.\ref{fig:reduction}. The
definitions of normal forms for values and computations are given in
Fig.\ref{fig:normal}. Subject reduction and strong normalization of
these rules can be proved using standard techniques.

@@

@@

@@

\end{document}

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\glob}{\mathop{\Box}}
\newcommand{\cat}{\mathbb{C}}
\newcommand{\catobj}{C}
\newcommand{\op}{\mathtt{op}}
\newcommand{\catop}{\cat^{\op}}
\newcommand{\id}[1]{\mathrm{id}_{#1}}
\newcommand{\setc}{\mathrm{Set}}
\newcommand{\psh}{\widehat{\cat}}
\newcommand{\terminal}{\top}
\newcommand{\initial}{\bot}
\newcommand{\lawc}{\mathrm{Law}}
\newcommand{\lawcop}{\lawc^{\mathtt{op}}}
\newcommand{\pshlawcop}{\widehat{\lawcop}}
\newcommand{\freef}{Free}
\newcommand{\constfname}{\Omega}
\newcommand{\constf}[1]{\constfname #1}
\newcommand{\yonedaname}{y}
\newcommand{\yoneda}[1]{y(#1)}
\newcommand{\homset}[3]{\mathrm{Hom}_{#1}(#2, #3)}
\newcommand{\arrow}[2]{{#2}^{#1}}
\newcommand{\sem}[1]{\llbracket #1 \rrbracket}
\newcommand{\reflectname}{\Uparrow}
\newcommand{\reflectraw}[1]{\mathop{\reflectname_{#1}}}
\newcommand{\reflect}[3]{\reflectraw{#1}(#2, #3)}
\newcommand{\reflectc}[2]{\mathop{\reflectname_{#1 #2}}}
\newcommand{\performraw}[2]{\mathop{\reflectname(#1(#2))}}
\newcommand{\perform}[5]{\performraw{#1}{#2}(#3, #4. #5)}
\newcommand{\reifyname}{\Downarrow}
\newcommand{\reifyraw}[1]{\mathop{\reifyname_{#1}}}
\newcommand{\reify}[3]{\reifyraw{#1}(#2.\,#3)}
\newcommand{\reifyc}[2]{\mathop{\reifyname_{#1 #2}}}
\newcommand{\outlname}{\mathrm{out}_L}
\newcommand{\outl}[1]{\mathop{\outlname} #1}
\newcommand{\outrname}{\mathrm{out}_R}
\newcommand{\outr}[1]{\mathop{\outrname} #1}
\newcommand{\mon}{\mathcal{T}}
\newcommand{\ind}[1]{W_{#1}}
\newcommand{\indintro}[4]{#2_{#1}(#3,\, #4)}
\newcommand{\retname}{\mathrm{ret}}
\newcommand{\ret}[2]{\mathop{\retname_{#1}}(#2)}
\newcommand{\indelim}[3]{\mathop{\mathrm{rec}_{#1}} #2 \,\{#3\}}
\newcommand{\indretcase}[2]{\retname(#1) \Rightarrow #2}
\newcommand{\indcase}[4]{#1(#2, #3) \Rightarrow #4}
\newcommand{\seq}{\,;\,}
\newcommand{\fst}{\mathop{\pi_1}}
\newcommand{\snd}{\mathop{\pi_2}}
\newcommand{\inl}[1]{\mathop{\mathrm{in}_L} #1}
\newcommand{\inr}[1]{\mathop{\mathrm{in}_R} #1}
\newcommand{\case}[5]{\mathop{\mathrm{case}} #1 \,\{\,\inl{#2} \Rightarrow #3 \seq \inr{#4} \Rightarrow #5 \,\}}
\newcommand{\abs}[3]{\mathop{\lambda}(#1 \types #2).\,#3}
\newcommand{\app}[2]{#1\,#2}
\newcommand{\types}{\mathrel{:}}
\newcommand{\gtypes}{\mathrel{:_{\Box}}}
\newcommand{\cempty}{\varepsilon}
\newcommand{\ccons}[2]{#1;\,#2}
\newcommand{\lbind}[3]{\ccons{#1}{#2\types#3}}
\newcommand{\gbind}[3]{\ccons{#1}{#2\gtypes#3}}
\newcommand{\cdiv}[1]{#1/\Box}
\newcommand{\bdiv}[1]{\cdiv{\mathop{!_{#1}}}}
\newcommand{\return}[1]{\mathop{\mathrm{return}} #1}
\newcommand{\letv}[3]{\mathop{\mathrm{let}} #1 = #2 \mathop{\mathrm{in}} #3}
\newcommand{\gbox}[1]{\mathop{\mathrm{box}} #1}
\newcommand{\gunbox}[3]{\mathop{\mathrm{let}} (\gbox #1) = #2 \mathop{\mathrm{in}} #3}
\newcommand{\grun}[1]{\mathop{\mathrm{run}}(#1)}
\newcommand{\sbst}[3]{#1[#2\backslash#3]}
\newcommand{\turnv}{\mathrel{\vdash_V}}
\newcommand{\turnc}{\mathrel{\vdash_C}}
\newcommand{\evmap}{\mathop{ev}}
\newcommand{\currymap}{\mathop{\lambda}}
\newcommand{\reducestov}{\rightsquigarrow_V}
\newcommand{\reducestoc}{\rightsquigarrow_C}
\newcommand{\mif}{\mathop{\mathrm{if}}}
\newcommand{\mthen}{\mathop{\mathrm{then}}}
\newcommand{\melse}{\mathop{\mathrm{else}}}
\newcommand{\embed}[1]{\mathopen{\lbag}#1\mathclose{\rbag}}
\newcommand{\ctxeq}{\cong_{\mathrm{ctx}}}

@@

We define the operational semantics as a pair of mutually defined
reduction relations $\reducestov$ and $\reducestoc$ for values and
computations respectively shown in Fig.\ref{fig:reduction}. The
definitions of normal forms for values and computations are given in
Fig.\ref{fig:normal}.

@@

We prove subject reduction for this semantics in
Appendix~\ref{apdx:subject-reduction} by straightforward induction on
the reduction relations.
\begin{lemma}
\label{lemma:subject-reduction}
  \begin{equation*}
    \Gamma \turnv v_1 \types \tau
    \wedge v_1 \reducestov v_2 \implies \Gamma \turnv v_2 \types \tau
  \end{equation*}
\end{lemma}

These reduction relations are strongly normalizing. We prove this via
Tait computability\cite{??} in
Appendix~\ref{apdx:strong-normalization}. This is made more involved by
the presence of infinitary inductive types, which requires us to
\todo{something something Knaster-Tarski}.
\begin{lemma}
\label{lemma:strong-normalization}
\todo{Find simplest formal description of strong normalization}
\end{lemma}

@@

\subsection{Properties of the denotational semantics}

The denotational semantics is sound:
\begin{lemma}
\label{lemma:soundness}
  \begin{equation*}
    \Gamma \turnv v_1 \types \tau
    \wedge v_1 \reducestov v_2
    \implies \sem{\Gamma \turnv v_1 \types \tau}
           = \sem{\Gamma \turnv v_2 \types \tau}
  \end{equation*}
\end{lemma}
and adequate for contextual equality $\ctxeq$:
\begin{lemma}
\label{lemma:adequacy}
  \begin{equation*}
    \sem{\turnv v_1 \types \tau} = \sem{\turnv v_2 \types \tau}
    \implies v_1 \ctxeq v_2
  \end{equation*}
\end{lemma}

We present the full semantics, precisely define contextual equality, and
prove these properties in Appendix~\ref{apdx:denotational}.

@@

\appendix
\section{Operational semantics}
\subsection{Subject reduction}
\label{apdx:subject-reduction}
\subsection{Strong normalization}
\label{apdx:strong-normalization}
\section{Denotational semantics}
\label{apdx:denotational}
\subsection{Soundness}
\subsection{Contextual equivalence}
\subsection{Compositionality}
\subsection{Adequacy}
\end{document}
